Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge \\ (Supplementary Material)

### 2025-02-08 paper/appendix.tex
Updates

We use Pillai's Trace because: 1) It uses all the canonical correlation values,
capturing the full extent of association, 2) Its interpretation is similar to
Pearson's correlation coefficient, i.e., $ 0 $ signifies no association and $ 1
$ signifies perfect linear relationship. We use a normalized version because
when comparing between pair of variables with different number of categories.
Our proposed measure has several desirable properties that make it well-suited
for our application:

\begin{enumerate}
	\item \textbf{Bounded: } The measure is bounded between $ 0 $ (no
		association) to $ 1 $ (perfect linear association), simplifying
		interpretation.
	% \item \textbf{Independent of Sample Size: } 
	\item \textbf{Invariant to the Dummy Encoding: } For categorical
		variables, residuals are computed using dummy encoding. Since
		canonical correlations identifies linear combinations across
		columns, this measure is invariant to the specific dummy
		encoding scheme used.
	\item \textbf{Equivalent to Partial Correlation for Continuous Variables: }
		When both $ X $ and $ Y $ are continuous variables, our measure
		is equivalent to the absolute value of partial correlation coefficient.
	\item \textbf{Equivalent to polychoric and polyserial correlation for
			ordinal and continuous variables: }
		Under the assumption that the ordinal variable is generated by
		discretizing an underlying gaussian continuous variable, this
		effect size is equivalent to polychoric and polyserial
		correlation. As both of them recover the Pearson correlation
		coefficient.
		\todo[inline]{Verify if this is correct}
	\item \textbf{Related to Cram\'er's V: } 
\end{enumerate}

In essence, our measure of conditional association extends existing metrics to mixed data, providing 
a unified measure that is interpretable.

>>>>

We use a normalized version of Pillai's Trace as it is bounded between $ 0 $
and $ 1 $ and easier to interpret.

% We use Pillai's Trace because: 1) It uses all the canonical correlation values,
% capturing the full extent of association, 2) Its interpretation is similar to
% Pearson's correlation coefficient, i.e., $ 0 $ signifies no association and $ 1
% $ signifies perfect linear relationship. We use a normalized version because
% when comparing between pair of variables with different number of categories.
% Our proposed measure has several desirable properties that make it well-suited
% for our application:

% \begin{enumerate}
% 	\item \textbf{Bounded: } The measure is bounded between $ 0 $ (no
% 		association) to $ 1 $ (perfect linear association), simplifying
% 		interpretation.
% 	% \item \textbf{Independent of Sample Size: } 
% 	\item \textbf{Invariant to the Dummy Encoding: } For categorical
% 		variables, residuals are computed using dummy encoding. Since
% 		canonical correlations identifies linear combinations across
% 		columns, this measure is invariant to the specific dummy
% 		encoding scheme used.
% 	\item \textbf{Equivalent to Partial Correlation for Continuous Variables: }
% 		When both $ X $ and $ Y $ are continuous variables, our measure
% 		is equivalent to the absolute value of partial correlation coefficient.
% 	\item \textbf{Equivalent to polychoric and polyserial correlation for
% 			ordinal and continuous variables: }
% 		Under the assumption that the ordinal variable is generated by
% 		discretizing an underlying gaussian continuous variable, this
% 		effect size is equivalent to polychoric and polyserial
% 		correlation. As both of them recover the Pearson correlation
% 		coefficient.
% 		\todo[inline]{Verify if this is correct}
% 	\item \textbf{Related to Cram\'er's V: } 
% \end{enumerate}
% 
% In essence, our measure of conditional association extends existing metrics to mixed data, providing 
% a unified measure that is interpretable.
